---
title: "Data Science HW#2"
author: "Sophia Miller"
date: "10/1/2019"
output: github_document
---

# Loading necessary packages

```{r load_packages}
library(tidyverse)
library(readxl)
```

# Problem 1
## Read and clean Mr. Trash Wheel sheet
#### *The code chunk below does the following:*

* *Specifies the relevant sheet in the Excel file*
* *Omits rows without dumpster-specific data*
* *Rounds number of sports balls to the nearest integer and converts result to an integer variable*

```{r read_trashwheel}
trashwheel_data = read_excel(
  path = "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N408",
  col_names = TRUE) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = as.integer(round(sports_balls, 0))
    )
``` 

## Read and clean precipitation data for 2017 and 2018
#### *The two code chunks below omit rows without precipitation data and add a `year` variable:*

```{r precip_2017}
precip_data_2017 = read_excel(
  path = "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2017 Precipitation",
  col_names = TRUE,
  skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017)
```


```{r precip_2018}
precip_data_2018 = read_excel(
  path = "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
  sheet = "2018 Precipitation",
  col_names = TRUE,
  skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018)
```

## Combining precipitation datasets and converting `month` to a character variable

```{r combine_precip}
precip_data = 
  full_join(precip_data_2017, precip_data_2018, by = NULL) %>%
 mutate(
    month = as.numeric(month),
    month = month.name[month]
    )
```


There are `r nrow(precip_data)` observations and `r ncol(precip_data)` variables in the precipitation dataset. Key variables in the precipitation dataset include month, year, and total precipitation. Total precipitation in 2018 was `r sum(pull(filter(precip_data, year == 2018), total))`. Total precipication in 2017 was `r sum(pull(filter(precip_data, year == 2017), total))`. 

There are in `r nrow(trashwheel_data)` observations and `r ncol(trashwheel_data)` variables in the trashwheel dataset. Key variables include dumpster number (`dumpster`), weight of litter collected (`weight_in_tons`), volume of litter collected (`volume_cubic_yards`), and numberof homes powered. For example, `r sum(pull(filter(trashwheel_data, year == 2019, month == "May"), weight_tons))` tons of trash was collected in May 2019. The dataset also contains information on the quantity of different types of litter collected in a dumpster. For example, the median number of sports balls in a dumpster in 2017 was `r median(pull(filter(trashwheel_data, year == 2017), sports_balls))` and the median number of plastic bottles in a dumpster in 2018 was `r median(pull(filter(trashwheel_data, year == 2018), plastic_bottles))`.

# Problem 2

## Read and clean pols-month data
#### *The code chunk below does the following:*

* *Separates the `mon` variable into `year`, `month`, and `day`*
* *Converts `month` to a character variable*
* *Creates a `president` variable that indicates if a president at a given year/month was a democrat or republican*
* *Removes the `prez_dem`, `prez_gop`, and `day` variables*

```{r pols_month}
pols_month_data = read_csv(file = "./data/pols-month.csv") %>%
janitor::clean_names() %>%
separate(mon, c("year","month","day")) %>%
  mutate(
    month = as.numeric(month),
    month = month.abb[month],
    president = prez_dem,
    president = recode(president, '1' = "dem", '0' = "gop"),
    year = as.numeric(year)) %>%
  select(-day, -prez_gop, -prez_dem)
```
## Read and clean snp data
#### *The code chunk below cleans the data in snp.csv similarly to `pols_month_data`, arranges the data in chronological by year and month, and makes `year` and `month` the leading columns.*

```{r snp}
snp_data = read_csv(file = "./data/snp.csv") %>%
janitor::clean_names() %>%
separate(date, c("month","day","year")) %>%
select(year, month, close, -day) %>%
  mutate(month = as.factor(month)) %>%
  arrange(year, month) %>%
mutate(
  month = month.abb[(as.numeric(month))],
  year = as.numeric(year))
```

## Read and clean unemployment data
#### *The code chunk below converts the data in unemployment.csv from wide to long format and mutates the `month` variable values to title case to match the `month` values in `snp_data` and `pols_month_data` so that they can be joined.*

```{r unemployment}
unemployment_data = read_csv(file = "./data/unemployment.csv") %>%
janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percent_unemployment"
    ) %>%
    mutate(
      year = as.numeric(year),
      month = str_to_title(month))
```
## Join `pols_month_data`, `snp_data`, and `unemployment_data`

```{r combine_pols_snp}
politics_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month")) %>% 
  left_join( unemployment_data, by = c("year", "month"))
```

The `pols_month_data` dataset contains information about the number of politicians who are democratic or republican at a given time point. The `snp_data` dataset contains information about the closing value of Standard & Poor's stock market index on a given date. The `unemployment_data` dataset contains information about the percentage of unemployment in a given month of an associated year. The dataset `politics_data`, which results from combining the three datasets, has `r nrow(politics_data)` rows and `r ncol(politics_data)` columns. The years range from 1947 to 2015. Key variables in the `politics` dataset include  `president`, which indicates if the president was a democrat or republican during a given year/month, `unemployment`, which indicates the percent unemployment during a given year/month, and `close`, which indicates the closing value of Standard & Poor's stock market during a given year/month. 

# Problem 3
## Read and clean baby name popularity data
#### *The code chunk below standardizes ethnicity names and the case of the `childs_first_name` variable`, and removes duplicate rows.*

```{r baby_names}
baby_name_data = 
  read_csv("./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(
    ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC"),
    childs_first_name = str_to_upper(childs_first_name)) %>%
  distinct(.keep_all = TRUE)
```
## Making a table that shows the popularity of the name Olivia among girls over time

```{r olivia_table}
baby_name_data %>%
  filter(childs_first_name == "OLIVIA", gender == "FEMALE") %>% 
  select(year_of_birth, ethnicity, rank) %>%
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  ) %>% 
  knitr::kable(format = 'pandoc', caption = "Popularity of Name Olivia among Female Babies by Ethnicity Over Time (2011-2016)")
```

## Making a table that shows the most popular name among boys over time

```{r popular_boys_names}
baby_name_data %>% 
  filter(rank == "1", gender == "MALE") %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>%
  knitr::kable(format = 'pandoc', caption = "Most Popular Male Baby Name by Ethnicity Over Time (2011-2016)")
```

## Making a scatterplot that plots the number of children with a given name vs. the popularity rank of that name

```{r scatterplot}
baby_subset = 
  filter(baby_name_data, gender == "MALE", ethnicity == "WHITE NON HISPANIC", year_of_birth == "2016" )

ggplot(baby_subset, aes(x = rank, y = count)) + geom_point()
```



